---
title: "p8105_hw2_lt2949"
output: github_document
date: "2023-10-01"
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```


# Problem1
Clean the data in `pols-month.csv` and replace the month number with month name. 
```{r}
#Clean the data in 'pols-month.csv'
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez"))
```

Clean the data in `snp.csv` with similar method.
```{r}
# Clean the data in 'snp.csv'
snp = 
  read_csv("./fivethirtyeight_datasets/snp.csv") |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  arrange(year, month) |>
  mutate(month = month.name[month]) |>
  select(year, month, close) 
```

We tidy the `unemployment` data so that it can be merged with the `pols` and `snp` datasets.
```{r}
# Clean the data in 'umemployment.csv'
unemployment = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

Merge the three datasets.
```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
```
The merged dataset is derived from the three previously sets, unified using 'year' and 'month' as criteria. Covering a spectrum of political, economic, and unemployment metrics, its coverage extends from 1947 through 2015. The `combined_data` set holds 822 records across 11 distinct columns. Key variables present in this set are: `year`, `month`, `gov_gop`, `sen_gop`, `rep_gop`, `gov_dem`, `sen_dem`, `rep_dem`, `president`, `close`, and `unemployment_rate`.


# Problem2
```{r}
# Import and clean Mr.Trash Wheel data
mr_trash_wheel <- 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 1, range = "A2:N586") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster))|>
  mutate(homes_powered = (weight_tons * 500) %/% 30)

# Import and clean Professor Trash Wheel data
prof_trash_wheel <- 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 2, range = "A2:M108") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster))|>
  mutate(homes_powered = (weight_tons * 500) %/% 30)

# Import and clean Gwynnda Trash Wheel data
gwy_trash_wheel <- 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 4, range = "A2:L157") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster))|>
  mutate(homes_powered = (weight_tons * 500) %/% 30)

# Add a new variable 
mr_trash_wheel <- mr_trash_wheel %>% mutate(wheel_name = "Mr. Trash Wheel")

prof_trash_wheel <- prof_trash_wheel %>% mutate(wheel_name = "Professor Trash Wheel")

gwy_trash_wheel <- gwy_trash_wheel %>% mutate(wheel_name = "Gwynnda Trash Wheel")

# Convert "year" to numeric
mr_trash_wheel$year <- as.numeric(mr_trash_wheel$year)
prof_trash_wheel$year <- as.numeric(prof_trash_wheel$year)
gwy_trash_wheel$year <- as.numeric(gwy_trash_wheel$year)

# Combine all datasets
trash_wheel_collection <- bind_rows(mr_trash_wheel, prof_trash_wheel, gwy_trash_wheel)
str(trash_wheel_collection)

```

The `trash_wheel_collection` dataset combines three separate datasets: `mr_trash_wheel`, `prof_trash_wheel`, and `gwy_trash_wheel`. It consists of `r nrow(trash_wheel_collection)` observations and `r ncol(trash_wheel_collection)` variables. Within the `trash_wheel_collection` dataset, there are variables detailing collection dates (i.e., `date`), a variable representing the total weight of the trash (i.e., `weight_tons`), variables denoting different types of trash (e.g., `plastic_bags`), and a derived data variable (i.e., `homes_powered`). According to the available data, Professor Trash Wheel collected a total weight of `r sum(prof_trash_wheel$weight_tons)` tons of trash. Additionally, in July 2021, Gwynnda collected `r sum(gwy_trash_wheel$cigarette_butts[gwy_trash_wheel$month == "July" & gwy_trash_wheel$year == 2021])` cigarette butts.


# Problem3
```{r}
# Import and clean the baseline
baseline <- 
  read.csv("data_mci/MCI_baseline.csv", skip = 1) |>
  janitor::clean_names() |>
  mutate(
      #Convert "sex" and "apoe4" to non-numeric
      sex = recode(sex, "1" = "male", "0" = "female"),
      apoe4 = recode(apoe4, "1" = "carrier", "0" = "non-carrier")
  ) |>
  rename("study_id" = "id")
  
# Remove participants who do not meet the stated inclusion criteria 
baseline_meet <- filter(baseline, is.na(age_at_onset) | age_at_onset == "." | current_age < age_at_onset)

# Filter the number of participants who developed MCI
mci_developed <- filter(baseline_meet, !is.na(age_at_onset) | age_at_onset != ".")

# Statistics process
women_apoe4_carriers <- sum(baseline$sex == "female" & baseline$apoe4 == "carrier")
total_women <- sum(baseline$sex == "female")
prop_women_carriers <- round((women_apoe4_carriers / total_women)*100, digits = 2)
```
After importing the data, we convert the variables `sex` and `apoe4` to more meaningful expressions. In the original baseline dataset, there are `r nrow(baseline)` observations. However, out of these, only `r nrow(baseline_meet)` participants met the specified requirements, as indicated by the number of observations in the `baseline_meet` dataset. Furthermore, `r nrow(mci_developed)` participants developed MCI. The average baseline age of the participants is `r round(mean(baseline_meet$current_age), digits = 2)`. In the study, the proportion of women who are carriers stands at `r prop_women_carriers`%.

```{r}
# Import and clean the longitudinally observed biomarker values dataset
amyloid <- 
  read.csv("data_mci/mci_amyloid.csv", skip = 1) |>
  janitor::clean_names()
  
amyloid_tidy <- pivot_longer(amyloid, baseline:time_8, names_to = "time", values_to = "years")
```
After using "pivot_longer" funtion, `amyloid_tidy` dataset has `r nrow(amyloid_tidy)` ovbservations and `r ncol(amyloid_tidy)` variables.

```{r}
# Combine datasets
combined_mci <- merge(baseline_meet, amyloid_tidy, by = "study_id")

# Check participants in only one of the datasets
only_baseline <- setdiff(baseline_meet$study_id, amyloid_tidy$study_id)
only_amyloid <- setdiff(amyloid_tidy$study_id, baseline_meet$study_id)

# Exporting the combined_mci to a CSV file
write.csv(combined_mci, "data_mci/combined_mci.csv", row.names = FALSE)
```
There are `r length(only_baseline)` participants only in baseline dataset, and there are `r length(only_amyloid)` participants only in amyloid dataset. The combined dataset has `r nrow(combined_mci)` observations and `r ncol(combined_mci)` variables.

