---
title: "p8105_hw2_lt2949"
output: github_document
date: "2023-10-01"
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```


# Problem1
Clean the data in `pols-month.csv` and replace the month number with month name. 
```{r}
#Clean the data in 'pols-month.csv'
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez"))
```

Clean the data in `snp.csv` with similar method.
```{r}
#Clean the data in 'snp.csv'
snp = 
  read_csv("./fivethirtyeight_datasets/snp.csv") |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  arrange(year, month) |>
  mutate(month = month.name[month]) |>
  select(year, month, close) 
```

We tidy the `unemployment` data so that it can be merged with the `pols` and `snp` datasets.
```{r}
#Clean the data in 'umemployment.csv'
unemployment = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

Merge the three datasets.
```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
```
The merged dataset is derived from the three previously sets, unified using 'year' and 'month' as criteria. Covering a spectrum of political, economic, and unemployment metrics, its coverage extends from 1947 through 2015. The `combined_data` set holds 822 records across 11 distinct columns. Key variables present in this set are: `year`, `month`, `gov_gop`, `sen_gop`, `rep_gop`, `gov_dem`, `sen_dem`, `rep_dem`, `president`, `close`, and `unemployment_rate`.


# Problem2
```{r}
#Import and clean Mr.Trash Wheel data
mr_trash_wheel <- 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 1, range = "A2:N586") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster))|>
  mutate(homes_powered = (weight_tons * 500) %/% 30)

#Import and clean Professor Trash Wheel data
prof_trash_wheel <- 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 2, range = "A2:M108") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster))|>
  mutate(homes_powered = (weight_tons * 500) %/% 30)

#Import and clean Gwynnda Trash Wheel data
gwy_trash_wheel <- 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 4, range = "A2:L157") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster))|>
  mutate(homes_powered = (weight_tons * 500) %/% 30)

#Add a new variable 
mr_trash_wheel <- mr_trash_wheel %>% mutate(wheel_name = "Mr. Trash Wheel")

prof_trash_wheel <- prof_trash_wheel %>% mutate(wheel_name = "Professor Trash Wheel")

gwy_trash_wheel <- gwy_trash_wheel %>% mutate(wheel_name = "Gwynnda Trash Wheel")

#Convert "year" to numeric
mr_trash_wheel$year <- as.numeric(mr_trash_wheel$year)
prof_trash_wheel$year <- as.numeric(prof_trash_wheel$year)
gwy_trash_wheel$year <- as.numeric(gwy_trash_wheel$year)

#Combine all datasets
trash_wheel_collection <- bind_rows(mr_trash_wheel, prof_trash_wheel, gwy_trash_wheel)
str(trash_wheel_collection)

```

The `trash_wheel_collection` dataset combines three separate datasets: `mr_trash_wheel`, `prof_trash_wheel`, and `gwy_trash_wheel`. It consists of `r nrow(trash_wheel_collection)` observations and `r ncol(trash_wheel_collection)` variables. Within the `trash_wheel_collection` dataset, there are variables detailing collection dates (i.e., `date`), a variable representing the total weight of the trash (i.e., `weight_tons`), variables denoting different types of trash (e.g., `plastic_bags`), and a derived data variable (i.e., `homes_powered`). According to the available data, Professor Trash Wheel collected a total weight of `r sum(prof_trash_wheel$weight_tons)` tons of trash. Additionally, in July 2021, Gwynnda collected `r sum(gwy_trash_wheel$cigarette_butts[gwy_trash_wheel$month == "July" & gwy_trash_wheel$year == 2021])` cigarette butts.


# Problem3
```{r}


```

